A planetary collision afterglow and transit of the resultant debris cloud

### 2023-10-12 src/tex/ms.tex
[showyourwork] overleaf sync

Figure \ref{fig:eclipse_overview} shows the optical light curves.
%
The start of the optical eclipse is seen around MJD 59350 in the $g'$ band observations, and returns to pre-eclipse levels by MJD 59850, giving a total eclipse duration of approximately 500 days.
%
The eclipse depth varies as a function of wavelength, which is discussed below.
%
The normalisation of the light curves for the ASAS-SN, ATLAS and ALLWISE photometry was by calculating the out of transit flux before MJD 58700.
%
The LCOGT and AAVSO photometry was determined by aligning the photometry of the eclipse with the ASAS-SN and ATLAS photometry on a per band basis.
%
By treating the linear changes in flux as due to the edges of large dust clouds crossing the disk of the star, and the largest amount of absorption within any of these segments as an estimate for the absorption of the cloud \citep[see equations 4.2 and 4.3 in ][]{Kennedy17}, we can determine a robust lower limit to the transverse velocity of 7.5 km\,s$^{-1}$ for the material moving in front of the star.
%
The method is illustrated in Extended Data Figure~\ref{fig:gradientconvert} and the results over the full optical transit are shown in the lower panel of Figure~\ref{fig:eclipse_overview}.
%
We convert the magnitude $M(t)$ at time $t$ to a normalised flux $f(t)=10^{(M(t)-M_0/-2.5)}$ where $M_0$ is the mean magnitude outside of the eclipse.
%
We visually determine turning points in the linearly increasing or decreasing photometric flux, fit straight lines to the selected points, and determine the flux gradient in units of day$^{-1}$.
%
A lower bound can be derived for the transverse velocity of the dust, $v$, by measuring the gradient of the light curve and determining what velocity a sharp edged and completely opaque occulter moving across the disk of the star would be needed to make the same gradient.
%
If the dust is on a circular orbit, it therefore has to be within 16\,au (equivalent to an orbital period of 63 years) around the star. The temperature of sub-micron dust grains at this distance would be approximately 200\,K.

@@

The SED in Extended Data Figure \ref{fig:sed} gives an estimate of the infrared flux that would arise from dust thrown off in the putative collision.
%
We assume that the collision occurred at of order 10\,au from the star, and hence that a dust temperature of 100\,K is reasonable.
%
Here 1\% of the mass of a 20\,$M_\oplus$ is assumed to be converted entirely into 0.1\,$\mu$m sized grains; this is an optimistic assumption given that the total mass thrown off in collisions is of order 1\% or less \citep{2012ApJ...745...79L}, and that this ejected mass is constituted of bodies of a range of sizes and all the mass is not present as dust.
%
The dust spectrum in Extended Data Figure \ref{fig:sed} is approximated as a blackbody multiplied by $0.01 \times (10 / \lambda)^{1.5}$ (with $\lambda$ in $\mu$m), based on the absorption efficiency for silicates \citep{1993ApJ...402..441L}.
%
Changing the grain size to 1\,$\mu$m yields a similar dust spectrum; while the dust emitting area is less, these grains emit more efficiently.
%
Thus, the mass in small grains detectable with ALMA is about 0.2\,$M_{\rm Earth}$.
%
The dust spectrum lies well below the WISE measurements, and just below the ALMA measurement.
%
Even in this optimistic case, thermal emission from dust thrown off in the collision is therefore not necessarily easily detected.
%
This difference is one of surface area; at 600\,pc thousands of square au of 100\,K dust is needed for a thermal detection with ALMA, but only a small fraction of a square au is needed to significantly dim a star (which has radius $\sim$0.01\,au).

>>>>

Figure \ref{fig:eclipse_overview} shows the optical light curves.
%
The start of the optical eclipse is seen around MJD 59350 in the $g'$ band observations, and returns to pre-eclipse levels by MJD 59850, giving a total eclipse duration of approximately 500 days.
%
The eclipse depth varies as a function of wavelength, which is discussed below.
%
The normalisation of the light curves for the ASAS-SN, ATLAS and ALLWISE photometry was by calculating the out of transit flux before MJD 58700.
%
The LCOGT and AAVSO photometry was determined by aligning the photometry of the eclipse with the ASAS-SN and ATLAS photometry on a per band basis.
%
By treating the linear changes in flux as due to the edges of large dust clouds crossing the disk of the star, and the largest amount of absorption within any of these segments as an estimate for the absorption of the cloud \citep[see equations 4.2 and 4.3 in ][]{Kennedy17}, we can determine a robust lower limit to the transverse velocity of 7.5 km\,s$^{-1}$ for the material moving in front of the star.
%
The method is illustrated in Extended Data Figure~\ref{fig:gradientconvert} and the results over the full optical transit are shown in the lower panel of Figure~\ref{fig:eclipse_overview}.
%
We convert the magnitude $M(t)$ at time $t$ to a normalised flux $f(t)=10^{\left(\left(M(t)-M_0\right)/-2.5\right)}$ where $M_0$ is the mean magnitude outside of the eclipse.
%
We visually determine turning points in the linearly increasing or decreasing photometric flux, fit straight lines to the selected points, and determine the flux gradient in units of day$^{-1}$.
%
A lower bound can be derived for the transverse velocity of the dust, $v$, by measuring the gradient of the light curve and determining what velocity a sharp edged and completely opaque occulter moving across the disk of the star would need to make the same gradient.
%
If the dust is on a circular orbit, it therefore has to be within 16\,au (equivalent to an orbital period of 63 years) around the star. The temperature of sub-micron dust grains at this distance would be in the range 100-200\,K.

@@

The SED in Extended Data Figure \ref{fig:sed} gives an estimate of the infrared flux that would arise from dust thrown off in the putative collision.
%
We assume that the collision occurred at of order 10\,au from the star, and hence that a dust temperature of 100\,K is reasonable.
%
Here 1\% of the mass of a \,$\mathrm{M_{Earth}}$ planet is assumed to be converted entirely into 0.1\,$\mu$m sized grains; this is an optimistic assumption given that the total mass thrown off in collisions is of order 1\% or less \citep{2012ApJ...745...79L}, and that this ejected mass is constituted of bodies of a range of sizes and all the mass is not present as dust.
%
The dust spectrum in Extended Data Figure \ref{fig:sed} is approximated as a blackbody multiplied by $0.01 \times (10 / \lambda)^{1.5}$ (with $\lambda$ in $\mu$m), based on the absorption efficiency for silicates \citep{1993ApJ...402..441L}.
%
Changing the grain size to 1\,$\mu$m yields a similar dust spectrum; while the dust emitting area is less, these grains emit more efficiently.
%
Thus, the mass in small grains detectable with ALMA is about 0.2\,$M_{\rm Earth}$.
%
The dust spectrum lies well below the WISE measurements, and just below the ALMA measurement.
%
Even in this optimistic case, thermal emission from dust thrown off in the collision is therefore not necessarily easily detected.
%
This difference is one of surface area; at 600\,pc thousands of square au of 100\,K dust is needed for a thermal detection with ALMA, but only a small fraction of a square au is needed to significantly dim a star (which has radius $\sim$0.005\,au).
